---
title: "Rclean"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Rclean}
  %\VignetteEncoding{UTF-8}
---

<!-- \VignetteEngine{knitr::knitr} -->
<!-- https://bookdown.org/yihui/rmarkdown/r-package-vignette.html -->

Written with research scientists in mind,
[Rclean](https://cran.r-project.org/web/packages/Rclean/)'s primary
function provides a simple way to isolate the minimal code you need to
produce specific results, such as a statistical table or a figure. By
analyzing the relationships among objects and functions, large and/or
complicated analytical scripts can be paired down to the
essentials. This can aid in debugging and re-factoring code and help
to make scientific projects more robust and easily shared.

# Quick-start Guide

You can install
[Rclean](https://cran.r-project.org/web/packages/Rclean/) from *CRAN*: 

```{r eval = FALSE}
install.packages("Rclean")
```

You can install the most up to date version using
[devtools](https://github.com/hadley/devtools):

```{r eval = FALSE}
install.packages("devtools")
devtools::install_github("MKLau/Rclean")
```

Once installed, per usual R practice, just load the *Rclean* package with:

```{r eval = TRUE}
library(Rclean)
```

```{r eval = TRUE, echo = FALSE, results = "hide"}
### Loads libraries and scripts
library(CodeDepends)
script <- system.file(
    "example", 
    "simple_script.R", 
    package = "Rclean")
script.long <- system.file(
    "example", 
    "long_script.R", 
    package = "Rclean")

```

[Rclean](https://cran.r-project.org/web/packages/Rclean/) usage is
simple. Just run the `clean` function with the file path to a script
as the input. We can use an example script that is included with the
package:

```{r eval = FALSE} 
script <- "example/simple_script.R"

```

Here's a quick look at the code:


```{r eval = TRUE}
readLines(script)

```

You can get a list of the variables found in an object with
`get_vars`. 

```{r eval = TRUE}
get_vars(script)

```

Sometimes for more complicated scripts, it can be helpful to see a
network graph showing the interdependencies of variables. `code_graph`
will produce a network diagram showing which lines of code produce or
use which variables (e.g. 1 -> "out"):

```{r eval = TRUE}
code_graph(script)
```

Now, we can pick the result we want to focus on for cleaning:


```{r eval = TRUE}
clean(script, "tab.15")
```

We can also select several variables at the same time:

```{r eval = TRUE}
my.vars <- c("tab.12", "tab.15")
clean(script, my.vars)
```

While just taking a look at the simplified code can be very helpful,
you can also save the code for later use or sharing (e.g. creating a
reproducible example for getting help) with `keep`:

```{r eval = FALSE}
my.code <- clean(script, my.vars)
keep(my.code, file = "results_tables.R")
```

If you would like to copy your code to your clipboard to copy-paste,
you can do that by not specifying a file path. You can now paste the
simplified as needed, such as into another script file or a help forum
thread.

```{r eval = FALSE}
keep(my.code)
```

# Some Thoughts on the Need for "Code Cleaning"

At it's root R is a statistical programming language. That it, it was
designed for use in analytical workflows. As such, the majority of the
R community is focused on producing code for idiosyncratic projects
that are results oriented. Also, R's design is intentionally at a
level that abstracts many aspects of programming that would otherwise
act as a barrier to entry for many users. This is good in that there
are many people who use R to their benefit with little to no formal
training in computer science or software engineering. However, these
same users are also frequently frustrated by code that is fragile,
buggy and complicated enough to quickly become obtuse even to
themselves in a very short amount of time. In addition, when scripts
take an extremely long time to execute, being able to reduce
unnecessary analyses can help increase computation efficiency.


More often then not, when someone is writing an R script, they're
intent on getting a set of results. This set of results is always a
subset of a much larger set of possible ways to explore a dataset, as
there are many statistical approaches and tests, let alone ways to
create visualizations and other representations of patterns in
data. This commonly leads to lengthy, complicated scripts from which
researchers manually subset results, but never refactor (i.e. reduce
to the final subset). In part, this is enabled by a lack of a proper
version control system, and in order to record their process and not
lose work, the entire process remains in a single or several
scripts. Although Rclean is not designed to fix the latter, it can
help with the former issue, once an appropriate versioning system is
adopted (e.g. git or subversion). 



# Example: Cleaning a Long Script

Conducting analyses is challenging in that it requires thinking about
multiple concepts at the same time. What did I measure? What analyses
are relevant to them? Do I need to transform the data? How do I go
about managing the data given how they were entered? What's the code
for the analysis I want to run? And so on. Data analysis can be messy
and complicated, so it's no wonder that code reflects this. And this
is a reason why having a way to isolate code based on variables can be
valuable.

The following is an example of a script that has some
complications. As you can see, although the script is not extremely
long, it's long enough to make it frustrating to visualize it in its
entirety and pick through it. 


```{R long-setup, echo = FALSE, eval = TRUE}

readLines(script.long)

```

So, let's say we've come to our script wanting to extract the code to
produce one of the results `fit.sqrt.A`, which is an analysis that is
relevant to some product. Not only do we want to double check the
results, we also want to use the code again for another purpose, such
as creating a plot of the patterns supported by the test. 

Manually tracing through our code for all the variables used in the
test and finding all of the lines that were used to prepare them for
the analysis would be annoying and difficult, especially given the
fact that we have used "x" as a prefix for multiple unrelated objects
in the script. Instead, we can easily do this automatically with
*Rclean*.

```{R long}

clean(script.long, "fit.sqrt.A")

```

As you can see, *Rclean* has picked through the tangled bits of code
and found the minimal set of lines relevant to our object of
interest. This code can now be visually inspected to adapt the
original code or ported to a new, "refactored" script. 

Related to the latter possibility, one note about comments. Although,
there is often very useful for even invaluable content in comments,
the `clean` function removes comments when isolating code. This is
partly due to the difficulty in determining the relevance of comments
to sections of code. Comments at the end of lines are typically
relevant to the line they are on, but not always, and comments
occupying their own lines usually refer to the following lines, but
this is also not necessarily the case. As `clean` depends on
determining relationships among lines based on variables and function
calls, it's not actually possible to do this automatically for a
general set of use cases. However, comments in the original code
remain untouched and can be used to inform the reduced code, and, as
the `clean` function is oriented toward isolating code based on a
specific result, the resulting code tends to naturally would tend
toward a comment along the lines of "This generates result C using
inputs A and B by conducting processing steps that do ...".


# Example: Functions from Scripts

Naturally following from the previous example, there are instances
where someone might want to create a function based on a specific
result. This might be the case for users of workflow systems, such as
*Drake*, which requires the creation of project specific
functions. We can do this with the following steps:

1. Decide on the object(s) we want.
2. Use `clean` to isolate the code to produce the object(s).
3. Determine the correct inputs and create arguments for them.
4. Create an object to be returned by the function.
5. Name the function.

For example, let's create a function to run the analysis that we
isolated in the previous example.



```{R tangle, echo = TRUE, eval = FALSE}

### Data Munging
### More Data Munging
### Make a plot
### Run another analysis
### Make Another Plot

```

# Behind the Scenes: How Rclean Works

## Getting the Graph of Functions and Objects with Provenance

[\\]: # ### Provenance
[\\]: #     ## 1 Why?
[\\]: #     ## 2 What is it?
[\\]: #     ## 3 Prospective and retrospective
[\\]: #     ## 4 CodeDepends
[\\]: #     ## 5 Other projects: point to End-to-end/provClean

# Results Orientation

[\\]: # ### How variables are isolated
[\\]: #     ## 1 Single variable selection
[\\]: #     ## 2 How multiple variables are isolated
[\\]: # ### More complicated script example
[\\]: # ### Applied to Drake

# Final Thoughts

[\\]: # Drake
[\\]: # Other Workflow Projects
