---
title: "Rclean"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Rclean}
  %\VignetteEncoding{UTF-8}
---

<!-- \VignetteEngine{knitr::knitr} -->
<!-- https://bookdown.org/yihui/rmarkdown/r-package-vignette.html -->

Written with research scientists in mind,
[Rclean](https://cran.r-project.org/web/packages/Rclean/)'s primary
function provides a simple way to isolate the minimal code you need to
produce specific results, such as a statistical table or a figure. By
analyzing the relationships among objects and functions, large and/or
complicated analytical scripts can be paired down to the
essentials. This can aid in debugging and re-factoring code and help
to make scientific projects more robust and easily shared.

# Quick-start Guide

You can install
[Rclean](https://cran.r-project.org/web/packages/Rclean/) from *CRAN*: 

```{r eval = FALSE}
install.packages("Rclean")
```

You can install the most up to date version using
[devtools](https://github.com/hadley/devtools):

```{r eval = FALSE}
install.packages("devtools")
devtools::install_github("MKLau/Rclean")
```

Once installed, per usual R practice, just load the *Rclean* package with:

```{r eval = TRUE}
library(Rclean)
```

```{r eval = TRUE, echo = FALSE, results = "hide"}
library(CodeDepends)
script <- system.file(
    "example", 
    "simple_script.R", 
    package = "Rclean")

```

[Rclean](https://cran.r-project.org/web/packages/Rclean/) usage is
simple. Just run the `clean` function with the file path to a script
as the input. We can use an example script that is included with the
package:

```{r eval = FALSE} 
script <- "example/simple_script.R"

```

Here's a quick look at the code:


```{r eval = TRUE}
readLines(script)

```

You can get a list of the variables found in an object with
`get_vars`. 

```{r eval = TRUE}
get_vars(script)

```

Sometimes for more complicated scripts, it can be helpful to see a
network graph showing the interdependencies of variables. `code_graph`
will produce a network diagram showing which lines of code produce or
use which variables (e.g. 1 -> "out"):

```{r eval = TRUE}
code_graph(script)
```

Now, we can pick the result we want to focus on for cleaning:


```{r eval = TRUE}
clean(script, "tab.15")
```

We can also select several variables at the same time:

```{r eval = TRUE}
my.vars <- c("tab.12", "tab.15")
clean(script, my.vars)
```

While just taking a look at the simplified code can be very helpful,
you can also save the code for later use or sharing (e.g. creating a
reproducible example for getting help) with `keep`:

```{r eval = FALSE}
my.code <- clean(script, my.vars)
keep(my.code, file = "results_tables.R")
```

If you would like to copy your code to your clipboard to copy-paste,
you can do that by not specifying a file path. You can now paste the
simplified as needed, such as into another script file or a help forum
thread.

```{r eval = FALSE}
keep(my.code)
```

# Some Thoughts on the Need for "Code Cleaning"

At it's root R is a statistical programming language. That it, it was
designed for use in analytical workflows. As such, the majority of the
R community is focused on producing code for idiosyncratic projects
that are results oriented. Also, R's design is intentionally at a
level that abstracts many aspects of programming that would otherwise
act as a barrier to entry for many users. This is good in that there
are many people who use R to their benefit with little to no formal
training in computer science or software engineering. However, these
same users are also frequently frustrated by code that is fragile,
buggy and complicated enough to quickly become obtuse even to
themselves in a very short amount of time. 


# Example: Cleaning a Complicated Script
# Example: Debugging a Broken Script

Code returns an error message, but the code is too messy and hard to
trace back to the error.

```{R debug, eval = FALSE}



```

# Example: Functions from Scripts

For example, here is a complicated, messy script:


```{R tangle, echo = TRUE, eval = FALSE}

### Data Munging
### More Data Munging
### Make a plot
### Run another analysis
### Make Another Plot

```

# Behind the Scenes: How Rclean Works

## Getting the Graph of Functions and Objects with Provenance

[\\]: # ### Provenance
[\\]: #     ## 1 Why?
[\\]: #     ## 2 What is it?
[\\]: #     ## 3 Prospective and retrospective
[\\]: #     ## 4 CodeDepends
[\\]: #     ## 5 Other projects: point to End-to-end/provClean

# Results Orientation

[\\]: # ### How variables are isolated
[\\]: #     ## 1 Single variable selection
[\\]: #     ## 2 How multiple variables are isolated
[\\]: # ### More complicated script example
[\\]: # ### Applied to Drake

# Final Thoughts

[\\]: # Drake
[\\]: # Other Workflow Projects
