---
title: "Rclean"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Rclean}
  %\VignetteEncoding{UTF-8}
---

<!-- \VignetteEngine{knitr::knitr} -->
<!-- https://bookdown.org/yihui/rmarkdown/r-package-vignette.html -->

# Introduction

Written with research scientists that are results oriented in mind,
[Rclean](https://cran.r-project.org/web/packages/Rclean/)'s primary
function provides a simple way to isolate the minimal code you need to
produce a specific result, such as a statistical table or a figure. By
focusing on specific results, large and/or complicated analytical
scripts can be paired down to the essentials and easily re-factored to
be more robust and easily shared.

# Quick-start Guide

## Install

You can install
[Rclean](https://cran.r-project.org/web/packages/Rclean/) from *CRAN*: 

```{r eval = FALSE}
install.packages("Rclean")
```

You can install the most up to date version easily with
[devtools](https://github.com/hadley/devtools):

```{r eval = FALSE}
install.packages("devtools")
devtools::install_github("MKLau/Rclean")
```

Once installed, per usual R practice, just load the *Rclean* package with:

```{r eval = TRUE}
library(Rclean)
```

```{r eval = TRUE, echo = FALSE, results = "hide"}
library(CodeDepends)
script <- system.file(
    "example", 
    "simple_script.R", 
    package = "Rclean")

```

[Rclean](https://cran.r-project.org/web/packages/Rclean/) usage is
simple. Have a script with code you want to clean saved to disk. Then,
just run the `clean` function with the path to the script as the
input. Here, we can use an example script that is included with the
package:

```{r eval = FALSE} 
script <- "example/simple_script.R"

```

Here's a quick look at the code:


```{r eval = TRUE}
readLines(script)

```

You can get a list of the variables found in an object with
`get_vars`. 

```{r eval = TRUE}
get_vars(script)

```

Sometimes for more complicated scripts, it can be helpful to see a
 network graph showing the interdependencies of
 variables. `code_graph` will produce a network diagram showing which
 lines of code produce or use which variables (e.g. 1 -> "out"):

```{r eval = TRUE}
code_graph(script)
```

Now, we can pick the result we want to focus on for cleaning:


```{r eval = TRUE}
clean(script, "tab.15")
```

We can also select several variables at the same time:

```{r eval = TRUE}
my.vars <- c("tab.12", "tab.15")
clean(script, my.vars)
```

While just taking a look at the simplified code can be very helpful,
you can also save the code for later use or sharing (e.g. creating a
reproducible example for getting help) with `keep`:

```{r eval = FALSE}
my.code <- clean(script, my.vars)
keep(my.code, file = "results_tables.R")
```

If you would like to copy your code to your clipboard, you can do that
by not specifying a file path. You can now paste the simplified as
needed.

```{r eval = FALSE}
keep(my.code)
```

# The Need for "Code Cleaning"

At it's roots R is a statistical programming language. R users are
largely focused on producing code for idiosyncratic projects. Also,
although R's low entry level (relative to other programming languages)
opens the door to scientists that do not have formal computer science
and software engineering training, this also leads to users into
situations where their code is often complicated, prone to failure and
hard to reproduce or reuse in different environments.

For example, here is a complicated, messy script:


```{R tangle, echo = TRUE, eval = FALSE}

### Data Munging

### More Data Munging

### Make a plot

### Run another analysis

### Make Another Plot


```

# Use-case: Refactoring



# Use-Case: Debugging

Code returns an error message, but the code is too messy and hard to
trace back to the error.

```{R debug, eval = FALSE}



```



# Behind the Scenes: Provenance


[\\]: # ### Provenance
[\\]: #     ## 1 Why?
[\\]: #     ## 2 What is it?
[\\]: #     ## 3 Prospective and retrospective
[\\]: #     ## 4 CodeDepends
[\\]: #     ## 5 Other projects: point to End-to-end/provClean

# Results Orientation

[\\]: # ### How variables are isolated
[\\]: #     ## 1 Single variable selection
[\\]: #     ## 2 How multiple variables are isolated
[\\]: # ### More complicated script example
[\\]: # ### Applied to Drake

# Contributing

This is an open-source project. If you would like to contribute to the
project, please check out [CONTRIBUTING.md](CONTRIBUTING.md).

Please note that the 'Rclean' project is released with a
[Contributor Code of Conduct](CODE_OF_CONDUCT.md).
By contributing to this project, you agree to abide by its terms.

